---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "README-"
)
```

# Building a Bioconductor package using RStudio

## Key resources

- Docker image: [Docker Hub](https://hub.docker.com/repository/docker/saskiafreytag/making_bioconductor_pkg)
- Workshop material: [pkgdown website](https://saskiafreytag.github.io/making_bioconductor_pkg)

## Workshop description

This workshop gives an introductory overview of how to build a Bioconductor package.

Workshop participants will learn what makes a Bioconductor package special, how to get started building one and how to disseminate your final product.
This workshop will feature introductory material, 'live' coding, and Q&A, all of which are adapted from the content below.

### Instructor

- [Saskia Freytag](https://twitter.com/trashystats) (saskia.freytag@perkins.uwa.edu.au)
- [Peter Hickey](https://peterhickey.org/) (hickey@wehi.edu.au)

### Pre-requisites

- Basic knowledge of R syntax.
- Github account!!!

### Workshop Participation

Students will build a small package which is ideally version controlled.
There will be a Q&A session in the second half of the workshop.

### Workshop goals and objectives

#### Learning goals

- Understand the purpose of a package
- Learn about the essential building blocks of a package
- Develop your first simple package
- Become familiar with version control

## Installation

This workshop uses Bioconductor version 3.12.
At the time of writing, this is the current 'devel' version of Bioconductor, which can be installed following [these instructions](https://www.bioconductor.org/developers/how-to/useDevel/).

You can then install the packages necessary for this workshop using the following:

```{r, eval = FALSE}
library(BiocManager)
install(c("usethis", "roxygen2", "devtools", "goodpractice", "BiocCheck"))
```

Alternatively, you can might like to use Docker to run the workshop in a container with R, all the necessary packages, and RStudio.
This can be done as follows:

- Log in to RStudio at [http://localhost:8787](http://localhost:8787) using username `rstudio` and password `welcome-to-bioc2020`. Note that on Windows you need to provide your localhost IP address like `http://191.163.92.108:8787/` - find it using `docker-machine ip default` in Docker's terminal.
- Run `browseVignettes(package = "Building a Bioconductor Package")`. Click on one of the links, "HTML", "source", "R code".
    - In case of `The requested page was not found` error, add `help/` to the URL right after the hostname, e.g., [http://localhost:8787/help/library/making_bioconductor_pkg/doc/Building_a_Bioconductor_Packages.html](http://localhost:8787/help/library/making_bioconductor_pkg/doc/Building_a_Bioconductor_Packages.html). This is a [known bug](https://github.com/rocker-org/rocker-versioned/issues/178).
